
# Get ebird hotspot data for India

```{r}
library (rebird)
# takes 6s or so
system.time (
    h <- ebirdhotspotlist (regionCode = "IN")
)
index <- which (!is.na (h$lat) &
                !is.na (h$lng) &
                !is.na (h$numSpeciesAllTime))
h <- h [index, ]
saveRDS (h, "hotspotsIN.Rds")
```

```{r}
h <- readRDS ("hotspotsIN.Rds")
library (dplyr)
h <- group_by (h, lat, lng, locId) %>%
    summarise (n = sum (numSpeciesAllTime),
               .groups = "rowwise") %>%
    rename (long = lng) %>%
    relocate (locId, long, lat, n)
h$count <- seq (nrow (h))
```

# Pre-process OSM data with osmium

```{r}
cmd <- "osmium tags-filter --overwrite -o india-hw.pbf india-latest.osm.pbf w/highway"           # nolint
system (cmd)
cmd <- paste0 ("osmium tags-filter --overwrite -o india-hw-filt.pbf ",
               "india-latest.osm.pbf ",
               "w/highway=primary,primary_link,secondary,secondary_link,",
               "service,tertiary,tertiary_link,trunk,trunk_link")
system (cmd)
cmd <- "osmium tags-filter --overwrite -o india-nat.pbf india-latest.osm.pbf wr/natural"          # nolint
system (cmd)
cmd <- "osmium tags-filter --overwrite -o india-for.pbf india-latest.osm.pbf wr/landuse=forest"          # nolint
system (cmd)
```

# Code to get polygons of natural areas

This also merges all contiguous polygons. Currently works with the largest 500
polygons only, not all quarter million or so.

```{r}
library (osmextract)
library (osmdata)
library (sf)
x <- oe_read ("india-nat.pbf", layer = "multipolygons")
x <- st_cast (x, "POLYGON")
x$area <- sf::st_area (sf::st_transform (x, 3857))
index <- order (x$area, decreasing = TRUE) [1:500]
xtest <- x [index, ]
y <- st_intersection (xtest)

x0 <- y [y$n.overlaps == 1, ]
x0$origins <- unlist (x0$origins)
x1 <- y [y$n.overlaps > 1, ]
x0 <- x0 [which (!x0$origins %in% unlist (x1$origins)), ]

o <- x1$origins
p <- lapply (o, function (i)
             st_cast (st_union (xtest [i, ]),
                      "POLYGON"))
p <- do.call (c, p)
g <- x0$geometry
if (!is.null (p))
    g <- c (g, p)
g <- st_sf (geometry = st_sfc (g, crs = 4326))
g$area <- sf::st_area (sf::st_transform (g$geometry, 3857))
g$index <- seq (nrow (g))
```

# Code to get polygons of roadless areas

This splits each natural polygon along any major roads, and calculates
statistics of resultant polygons. The input, `p`, is one row of an `sf`
"POLYGON" `data.frame` with an `index` column.

```{r}
roadless_area <- function (p) {

    a_nat <- sf::st_area (sf::st_transform (p, 3857))

    null_obj <- c (a_rd_med = a_nat,
                   a_rd_max = a_nat,
                   a_nat = a_nat)

    f <- "polygon.poly"
    write.table (st_coordinates (p) [, 1:2],
                 file = f, sep = "    ",
                 row.names = FALSE, col.names = FALSE)
    x <- c ("polygon",
            "first_polygon",
            readLines (f),
            "END",
            "END")
    writeLines (x, f)

    f_hw <- sprintf ("data/india-hw%06d.osm", p$index)

    if (!file.exists (f_hw)) {

        cmd <- paste0 ("osmium extract -p polygon.poly india-hw-filt.pbf ",
                       "--overwrite -o ", f_hw)
        chk <- system (cmd, ignore.stdout = TRUE)
        count <- 0
        while (chk == 1 & count < 10) {
            chk <- system (cmd, ignore.stdout = TRUE)
            count <- count + 1
        }
    }

    dat <- osmdata_sf (doc = f_hw, quiet = TRUE) %>%
        osm_poly2line ()
    xl <- dat$osm_lines # nolint

    if (is.null (xl))
        return (null_obj)

    sp <- lwgeom::st_split (p, xl)
    sp <- sf::st_collection_extract (sp, "POLYGON")

    a <- sf::st_area (sf::st_transform (sp, 3857))

    a_rd_med <- median (a)
    a_rd_max <- max (a)

    return (c (a_rd_med = a_rd_med,
               a_rd_max = a_rd_max,
               a_nat = a_nat))
}
```

The actual calculation, and joining of these statistics with hotspot
statistics, where the latter are associated with statistics of containing
polygons.

```{r}
a <- pbapply::pblapply (seq (nrow (g)), function (i)
                        roadless_area (g [i, ]))
saveRDS (a, "a.Rds")
```
```{r}
a <- do.call (rbind, readRDS ("a.Rds")) / 1e6
#a <- roadless_area (g [5, ])
g <- g [seq (nrow (a)), ]
g <- cbind (g, a)
```

# Join polygon data to hotspot counts

The final step here involves removing the few very largest polygons.

```{r}
xy <- st_sfc (st_multipoint (as.matrix (h [, 2:3])),
              crs = 4326)
xy <- st_sf (locId = h$locId,
             n = h$n,
             geometry = st_cast (xy, "POINT"))
xy <- st_join (xy, g, join = st_within)
xy <- na.omit (xy)

index <- which (xy$a_nat < 4000)
message ("Removing ", nrow (xy) - length (index), " / ", nrow (xy), " largest polygons")
xy <- xy [index, ]
```

# Results


```{r}
library (ggplot2)
ggplot (xy, aes (x = a_rd_max, y = n)) +
    geom_point (pch = 19, col = "orange") +
    geom_smooth (method = "loess", formula = "y ~ x")
mod <- lm (n ~ a_nat + a_rd_med + a_rd_max, data = xy)
summary (mod)
mod <- lm (n ~ a_nat + a_rd_max, data = xy)
summary (mod)
```

